To conclude the first semester of our class with a bang, we will review all of the major categories we covered so far. Since there is not much space on this document, I recommend doing all work on a separate sheet of paper (and you can put your answers there too).
	\subsection{Counting}
		We did most of our counting through the AoPS videos. These videos are very helpful (and entertaining!) so they should have taught you some good counting skills. Let's take a look at a few of them:
		\begin{enumerate}
			\item How many numbers are in the set $\{-9, -7, \cdots, 199\}$
			\item How many multiples of 7 are there between 183 and 937?
			\item If I have 4 salads and 3 meats, how many meals can I make of 2 salads and 2 meats if order does not matter?
			\item If there are 8 horses in a race, how many ways are there for them to finish $\nth{1}$, $\nth{2}$ and $\nth{3}$? 
			\item How many ways are there to arrange the letters of ``DISNEY''?
			\item 5!
			\item 6!
			\item 7!
			\item 8!
			\item 9!
			\item 10!
			\item ${874 \choose 0}$ 
			\item ${959 \choose 1}$ 
			\item ${9 \choose 7}$ 
			\item ${19 \choose 3}$ 
			\item ${9999 \choose 9998}$ 
			\item How many ways can I walk from (0, 0) to (6, 7) on a grid?
			\item With the same $6 \times 7$ grid, how many of these paths go through (3, 4)?
			\item With the same $6 \times 7$ grid, how many of these paths go through (3, 4) and (6, 5)?
			\item How many ways can I walk along the edges of a 3x3x3 cube? (I can go through the center)
			\item How many ways can I walk along the surface of a 2x2x2 grid (I \textbf{cannot} go into the center)?
		\end{enumerate}
	\subsection{Number Theory}
		 The bulk of what we covered this year was in Number Theory---these techniques help you understand numbers better so you can work with math more proficiently.
		\subsubsection{Factors}
			\begin{enumerate}
				\item What number do I have to put in 482\_8 to make it divisible by 3 and 4?
				\item Given 180:
				\begin{enumerate}
					\item How many positive factors does it have?
					\item How many even factors does it have?
					\item How many odd factors does it have?
					\item How many perfect square factors does it have?
					\item What are the sum of the factors of it?
					\item Do (a)-(e) for 108, 576, 1296
				\end{enumerate}
				\item What is the largest power of 2 that divides 45!?
				\item What is the largest power of 4 that divides 45!?
				\item What is the largest power of 5 that divides 45!?
				\item Completely prime factorize 20!
				\item Why do perfect squares always have an odd number of factors?
				\item Is 0 divisible by anything? Everything?
			\end{enumerate}
		\subsubsection{Modular Arithmetic}
			\begin{enumerate}
				\item What is the last digit of $2^{2842}$?
				\item What is the last digit of $3^{1958^{1842}}$
				\item What is the last digit of $1382^{2842}$
				\item What is the last digit of $\sum_{i=1}^\infty i! = 1!+ 2!+3!+4!+5!+\cdots$
				\item What is the last digit of $1928999936!$ ?
				\item What is the remainder when $23+49+82-71$ is divided by 8?
				\item What is the remainder when $2938493 + 17329 + 4889382$ is divided by 4?
				\item What is the remainder when $3\times 8 \times 11$ is divided by 5?
				\item What is the remainder when $1248963 \times 242448481 \times 964836123 \times 24363636967$ is divided by 12?
				\item What is the smallest positive integer that has a remainder of 3 when divided by 5 and a remainder of 2 when divided by 17?
				\item What is the smallest positive integer that has a remainder of 2 when divided by 4, 5, 6, 7, and 8?
				\item What is the smallest positive integer that has a remainder of 4 when divided by 5, 7 when divided by 8, 12 when divided by 13 and 18 when divided by 19?
			\end{enumerate}
		\subsubsection{Sense}
			You've already seen very extreme cases that number sense can deal with. This section has problems that you will \textbf{definitely} see more often than something like $25 \times 18378275208$ but also has the extreme cases anyway for review.
			\begin{enumerate}
				\item $5 \times 135678992$
				\item $5 \times 48527888$
				\item $25 \times 96385264$
				\item $25 \times 123456789$
				\item $15 \times 32$
				\item $25 \times 45$
				\item $11 \times 19$
				\item $34 \times 36$
				\item $22 \times 41$
				\item $17 \times 71$
				\item $13 \times 34$
				\item List all all the squares from $1$ to $40$
				\item $85^2$
				\item $285^2$
				\item $65^4$
				\item $3 \times 142857$
				\item $\frac{184}{7}$ in decimal form
				\item $13 \times 17$
				\item $22 \times 34$
				\item $92 \times 98 \times 3$
				\item $37 \times 43$
				\item $98 \times 107$
			\end{enumerate}
	\subsection{Algebra}
		\begin{enumerate}
			\item If kid tickets cost \$2 and adult tickets cost \$3, how many kids and adults are there in a group of 86 people that spends \$244 total?
			\item $(a+b)^2 = $
			\item $(a-b)^2 = $
			\item $(a+b)^3 = $
			\item $(a+b)^3 = $
			\item $(a+b)^4 = $
			\item $(a+b)^4 = $
			\item Do you think the coefficients of the expansions above look like Pascal's triangle? If so, try to figure out what $(a+b)^5$ looks like and figure out in general what $(a+b)^n$ will look like
			\item $a^2-b^2 = $
			\item $a^3+b^3 = $
			\item $a^3-b^3 = $
			\item What is the largest prime factor of $82^{2}-62^{2}$?
			\item $\frac{8x}{7} + 8 = -19$
			\item $5x^{2} = 80$
			\item $\sqrt{9620}$
			\item $(\sqrt{8})^{\frac{25}{6}}$
			\item $\frac{1-\sqrt{15}}{\sqrt{5}-2}$ 
			\item $\frac{\sqrt{7}+\sqrt{3}}{\sqrt{18}-\sqrt{14}}$
			\item Estimate the value of $x$, if $\frac{1}{x} +1 = \sqrt[3]{3}$
			\item Compute $\sum\limits_{i=1}^{15} i$ \par
			\item Compute $\sum\limits_{i=8}^{28} 5$ \par
			\item Compute $\sum\limits_{i=1}^{\infty} \frac{1+3^{k}}{4^{k}}$ \par
			\item Compute $\prod\limits_{i=1}^{10} \frac{i}{i+2}$
			\item Estimate $0.99^{1839}$
			\item Estimate $1.002^{2752}$

		\end{enumerate}
			\subsection{Geometry}
		Mostly areas you should know for geometry---and of course the Pythagorean theorem. Since we did not get through much, I will not include many problems here.
		\begin{enumerate}
			\item What is the area of a square with side length $8\sqrt{2}$?
			\item What is the area of an equilateral triangle with side length $2-\sqrt{3}$
			\item A right triangle with hypotenuse 13 and one of its legs 5
			\item Area of a circle with diameter 5
			\item Area of a trapezoid with bases 3 and 4, and height 8
			\item Area of a rectangle with side lengths 14 and 3
			\item What is the hypotenuse of a triangle with legs 18 and 34?
			\item What is the area of an equilateral triangle inscribed inside a circle of radius 4?
			\item Find the distance from (-1, 8) to (9, 4)
			\item The measure of an interior angle of a regular 18-gon
			\item A cylindrical tank with height 17 m and radius 8 m is filled at the rate of 37 L/hour. The tank is being drained at 22 L/min. How long will it take to fill the tank? Note: there are 1000 liters per cubic meter
			\item A toy house consists of a 4 in cube with a square pyramid of height 2 in on top. What is the total volume of this house?
		\end{enumerate}